Seasonality in governing dynamics at submesoscales

### 2016-06-10 writeup/rocha_etal.tex
Update figure 1

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.6\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.65\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

\begin{acknowledgments}
We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NASA and NSF.
\end{acknowledgments}

@@

\subsection*{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

Figure \ref{figA1} shows KE frequency spectra for the model and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods > 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermedia frequencies. At 40 m, the model and observations
have consistent spectra at periods larger than about 6 h. The data from moorings
are more energetic at higher frequencies and tidal harmonics are not observed,
except for the first two harmonics that are apparent in the KESS data at 40 m.
Three reasons for the lack of tidal harmonics in the mooring data are: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full spectrum
of ultraviolet catastrophic wave-wave interactions; 3) The model is not forced at
frequencies higher that 6h.

@@

>>>>

\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.6\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.65\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. Colors represent the topography and black lines are absolute
          dynamic topography contours every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

\begin{acknowledgments}
We thank the MITgcm community and our colleagues at the NASA Advanced
Supercomputing (NAS) Division for their awesome support.
This research was funded by NASA and NSF.
The altimeter products were produced by Ssalto/Duacs
and distributed by Aviso, with support from Cnes (\texttt{http://www.aviso.altimetry.fr/duacs/})
\end{acknowledgments}

@@

\subsection*{How well does the LLC simulations captures high-frequency modes?}
To assess how well the LLC simulations represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
data are more energetic at intermedia frequencies. At 40 m, the LLC simulations and observations
have consistent spectra at periods larger than about 6 h. The data from moorings
are more energetic at higher frequencies,
except for the first two harmonics that are apparent in the KESS data at 40 m.
There are at least three reasons for the slightly more energetic super-inertial
variability in the observations: 1) The high-
frequency variability is dominated by high-frequency noise due to mooring vertical
excursions; 2) The model does not have enough resolution to resolve the full internal
wave spectrum; 3) The model is not forced at frequencies higher that 6h.

@@

Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
climatology in April and October. Clearly, the the changes to the stratification
are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
seasonality of upper-ocean stratification yields a strong seasonality in the
near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
In particular, the surface amplitude is much large in Summer. Thus the projection
of baroclinic tides on the surface may vary significantly seasonally.

\textcolor{red}{Perhaps include a model with a slab over a slowly varying
pycnocline. The model is simple, but the solution is a bit messy and may
scare GRL readers...}.
